A Type System for Safe and Scalable Parallelism using Logical Regions

### 2012-07-10 doc/popl2013/intro.tex

\lstset{
  captionpos=b,
  language=C++,
  basicstyle=\scriptsize,
  numbers=left,
  numberstyle=\tiny,
  columns=fullflexible,
  stepnumber=1,
  escapechar=\#,
  keepspaces=true,
  literate={<}{{$\langle$}}1 {>}{{$\rangle$}}1,
  morekeywords={region,coloring,partition,spawn,disjoint,aliased},
  deletekeywords=float,
}
\begin{lstlisting}[float={t},label={lst:code_ex},caption={Circuit simulation.}]
struct Node { float voltage, new_charge, capacitance; };
struct Wire<rn> { Node@rn in_node, out_node; float current, ... ; };
struct Circuit { region   r_all_nodes; /* contains all nodes for the circuit */
                 region   r_all_wires; /* contains all circuit wires */ };
struct CircuitPiece {
  region  rn_pvt, rn_shr, rn_ghost; /* private, shared, ghost node regions */
  region  rw_pvt;                   /* private wires region */ };

void simulate_circuit(Circuit c, float dt) : RWE(c.r_all_nodes, c.r_all_wires)
{
  // The construction of the colorings is not shown.  The colorings wire_owner_map,
  // node_owner_map, and node_neighbor_map have MAX_PIECES colors 
  // 0..MAX_PIECES #$-$# 1. The coloring node_sharing map has two colors 0 and 1.
  //
  // Partition of wires into MAX_PIECES pieces
  partition<disjoint> p_wires = c.r_all_wires.partition(wire_owner_map); 
  // Partition nodes into two parts for all-private vs. all-shared
  partition<disjoint> p_nodes_pvs = c.r_all_nodes.partition(node_sharing map);

  // Partition all-private into MAX_PIECES disjoint circuit pieces
  partition<disjoint> p_pvt_nodes = p_nodes_pvs[0].partition(node_owner_map);
  // Partition all-shared into MAX_PIECES disjoint circuit pieces
  partition<disjoint> p_shr_nodes = p_nodes_pvs[1].partition(node_owner_map);
  // Partition all-shared into MAX_PIECES ghost regions, which may be aliased
  partition<aliased> p_ghost_nodes = p_nodes_pvs[1].partition(node_neighbor_map);

  CircuitPiece pieces[MAX_PIECES];
  for(i = 0; i #$<$# MAX_PIECES; i++) 
    pieces[i] = { rn_pvt: p_pvt_nodes[i], rn_shr: p_shr_nodes[i],
                  rn_ghost: p_ghost_nodes[i], rw_pvt: p_wires[i] };
  for (t = 0; t #$<$# TIME_STEPS; t++) {
    for (i = 0; i #$<$# MAX_PIECES; i++) calc_new_currents(pieces[i]);
    for (i = 0; i #$<$# MAX_PIECES; i++) distribute_charge(pieces[i], dt);
    for (i = 0; i #$<$# MAX_PIECES; i++) update_voltages(pieces[i]);
  }
}

void calc_new_currents(CircuitPiece piece):
        RWE(piece.rw_pvt), ROE(piece.rn_pvt, piece.rn_shr, piece.rn_ghost) {
  foreach(w : piece.rw_pvt)
    w#$\rightarrow$#current = (w#$\rightarrow$#in_node#$\rightarrow$#voltage - w#$\rightarrow$#out_node#$\rightarrow$#voltage) / w#$\rightarrow$#resistance;
}

void distribute_charge(CircuitPiece piece, float dt):
        ROE(piece.rw_pvt), RdA(piece.rn_pvt, piece.rn_shr, piece.rn_ghost) {
  foreach(w : piece.rw_pvt) {
    w#$\rightarrow$#in_node#$\rightarrow$#new_charge += -dt * w#$\rightarrow$#current;
    w#$\rightarrow$#out_node#$\rightarrow$#new_charge +=  dt * w#$\rightarrow$#current;
  }
}

void update_voltages(CircuitPiece piece): RWE(piece.rn_pvt, piece.rn_shr) {
  foreach(n : piece.rn_pvt, piece.rn_shr) {
    n#$\rightarrow$#voltage += n#$\rightarrow$#new_charge / n#$\rightarrow$#capacitance;
    n#$\rightarrow$#new_charge = 0;
  }
}
\end{lstlisting}

>>>>

\begin{lstlisting}[float={t},label={lst:code_ex},caption={Circuit simulation.}]
type CircuitNode        = < float, --voltage
                            int >
type CircuitWire<rn,rg>  = < CircuitNode@rn,
                             CircuitNode@(rn,rg),
                             float> -- current

type node_list<rl,rn>       = < CircuitNode@rn, node_list<rl,rn>@rl >
type wire_list<rl,rw,rn,rg> = < CircuitWire<rn,rg>@rw, wire_list<rl,rw,rn,rg>@rl >

type CircuitPiece<rl,rw,rn> = rr[rpw,rpn,rg]
        < wire_list<rl,rpw,rpn,rg>@rl,
          node_list<rl,rpn>@rl >         
          where rpn <= rn and rg <= rn and rpw <= rw and
                rpn * rg and rn * rw and rl * rn and rl * rw

type piece_list<rl,rp,rw,rn> = < CircuitPiece<rl,rw,rn>@rp, piece_list<rl,rp,rw,rn>@rl >

function color_circuit[rn,rw,rl] ( all_nodes : node_list<rl,rn>@rl,
                               all_wires : wire_list<rl,rw,rn>@rl ), reads(rn,rw,rl) : 
                               <coloring(rn), coloring(rn), coloring(rw)> =  
  -- Invoke programmer chosen coloring algorithm (e.g. METIS)

function build_lists[rl,rw,rn,rpw,rpn,rg] ( all_nodes : node_list<rl,rn>@rl, all_wires : wire_list<rl,rw,rn>@rl,
                                 owned_coloring : coloring(rn), ghost_coloring : coloring(rn),
                                 wires_coloring : coloring(rw), c : int), reads(rn,rw,rl), writes(rl) :
                                 < wire_list<rl,rpw,rpn,rg>@rl, node_list<rl,rpn>@rl > = 
  -- Construct node and wire lists for each piece based on coloring

function build_piece_list[rl,rp,rw,rn] ( one: CircuitPiece<rl,rw,rn>, two : CircuitPiece<rl,rw,rn>,
                                 three: CircuitPiece<rl,rw,rn>, four : CircuitPiece<rl,rw,rn> ),
                                 reads(rl,rp,rw,rn), writes(rl,rp) : piece_list<rl,rp,rw,rn>@rl =
  -- Build a list of the pieces

function simulate_circuit[rl,rw,rn] ( all_nodes : node_list<rl,rn>@rl,
                                  all_wires : wire_list<rl,rw,rn,rn>@rl, steps : int ),
                          reads(rn,rw,rl), writes(rn,rw,rl) : bool = 
  let pc : <coloring(rn),<coloring(rn),coloring(rw)> >
                      = color_circuit[rn,rw,rl](all_nodes,all_wires) in
  partition rn using pc.1 as rn0,rn1,rn2,rn3 in
  partition rn using pc.2 as rg0,rg1,rg2,rg3 in
  partition rw using pc.3 as rw0,rw1,rw2,rw3 in
  let piece0_lists : < wire_list<rl,rw0,rn0,rg0>@rl, node_list<rl,rn0>@rl > = 
           build_lists[rl,rw,rn,rw0,rn0,rg0](all_nodes,all_wires,pc.1,pc.2.1,pc.2.2,0) in
  let piece0 : CircuitPiece<rl,rw,rn> = pack piece0_lists as CircuitPiece<rl,rw,rn>[rw0,rn0,rg0] in
  let piece1_lists : < wire_list<rl,rw1,rn1,rg1>@rl, node_list<rl,rn1>@rl > = 
           build_lists[rl,rw,rn,rw1,rn1,rg1](all_nodes,all_wires,pc.1,pc.2.1,pc.2.2,1) in
  let piece1 : CircuitPiece<rl,rw,rn> = pack piece1_lists as CircuitPiece<rl,rw,rn>[rw1,rn1,rg1] in
  let piece2_lists : < wire_list<rl,rw2,rn2,rg2>@rl, node_list<rl,rn2>@rl > = 
           build_lists[rl,rw,rn,rw2,rn2,rg2](all_nodes,all_wires,pc.1,pc.2.1,pc.2.2,2) in
  let piece2 : CircuitPiece<rl,rw,rn> = pack piece2_lists as CircuitPiece<rl,rw,rn>[rw2,rn2,rg2] in
  let piece3_lists : < wire_list<rl,rw3,rn3,rg3>@rl, node_list<rl,rn3>@rl > = 
           build_lists[rl,rw,rn,rw3,rn3,rg3](all_nodes,all_wires,pc.1,pc.2.1,pc.2.2,3) in
  let piece3 : CircuitPiece<rl,rw,rn> = pack piece3_lists as CircuitPiece<rl,rw,rn>[rw3,rn3,rg3] in
  let all_pieces : pieces_list<rl,rl,rw,rn> = build_piece_list[rl,rl,rw,rn](piece0,piece1,piece2,piece3) in
      execute_time_step[rl,rl,rw,rn](all_pieces,steps)

function execute_time_step[rl,rp,rw,rn] ( all_pieces : pieces_list<rl,rp,rw,rn>@rl ),
                                          reads(rl,rp,rw,rn), writes(rw,rn) : bool =
  if steps < 1 then true else
  let all_pieces' : pieces_list<rl,rp,rw,rn>@rl = map calc_new_currents[rl,rw,rn] all_pieces in
  let all_pieces'' : pieces_list<rl,rp,rw,rn>@rl = map distribute_charge[rl,rw,rn] all_pieces' in
  let all_pieces''' : pieces_list<rl,rp,rw,rn>@rl = map update_voltage[rl,rw,rn] all_pieces'' in
      execute_time_step[rl,rp,rw,rn](all_pieces''',steps-1)

function calc_new_currents[rl,rw,rn] ( p : CircuitPiece<rl,rw,rn> ), 
                                      reads(rl,rw,rn), writes(rw) : CircuitPiece<rl,rw,rn> =
  -- Calculate new currents for each wire

function distribute_charge[rl,rw,rn] ( p : CircuitPiece<rl,rw,rn> ), 
                      reads(rl,rw,rn), reduces(reduce_charge,rn), atomic(rn) : CircuitPiece<rl,rw,rn> =
  -- Distribute charge from each wire to all nodes

function update_voltage[rl,rw,rn] ( piece : CircuitPiece<rl,rw,rn> ), reads(rl,rw,rn), writes(rn) : bool = 
  -- Update voltage for all our owned nodes

\end{lstlisting}
